We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-02-15 paper/paper.tex
bib, y otras correcciones

\begin{abstract}
We use numerical simulations to build an explicit probability
distribution for the asphericity in the satellite distribution around
galaxies similar to the Local Group (LG) in the Lambda Cold Dark
Matter (LCDM) paradigm. 
This allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that fully resemble the LG.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-Dark and ELVIS. 
Detailed results differ among the simulations suggesting a strong
influence of the typical DM halo mass in the different pair samples.  
However, there are three common trends in these simulations. 
First, at most of $2\%$ the pairs are expected to have satellite
distributions with the same asphericity as the LG; second,
between $27\%$ to $56\%$ of the pairs have a halo with a satellite
distribution as aspherical as in M31; and third, at most $3\%$ of the
pairs have a satellite distribution as planar as in the MW. 
The brightest satellites in M31 galaxy turn out to have a rather
typical distribution in the LCDM context, while the MW is atypical in
this respect.  
These results place the LG at the level of a $3\sigma$ outlier
in the LCDM paradigm. 
The approach presented here facilitates the comparison between
different numerical setups and choices to select halo pairs in simulations.
Quantifying the probability density for the satellite distribuition
asphericity as a function of halo mass and large scale environment 
will be a necessary step to understand the origin of the LG
atipicallity in the LCDM context as presented here.  
\end{abstract}

@@

The spatial distribution of satellite galaxies around our Milky Way
(MW) and the M31 galaxy provides a stringent test for structure
formation theories in a explicit cosmological context.  
The presence of a Vast Polar Structure of satellites around the Milky
Way has been stablished in the last decade.
This structure is present in the $11$ brightest "classical" dwarf
galaxies and has been confirmed as fainter satellites have been
detected. 
In the case of M31,  observational studies have found a planar satellite
distribution in a subset of $15$ satellites out of the homogeneous
parent sample of $27$ satellites. 
However, taking into account the $11$ brightest satellites (as a
comparison with the MW) there isn't evidnce for any planar structure.

The question whether this kind of planar configurations are atypical
in the observable Universe cannot be properly addressed given the
current observational limitations.
Comparisons against observations in large galaxy surveys have only
been able to constrain that $5\%$ of the Milky-Way like galaxies in
the Universe have two bright satellite galaxies as is the case of the
Milky Way and its Large and Small Magellanic Clouds. 
A planar structure ouside the Local Group (LG) has only been confirmed
in Centaurus A \citep{muller2018whirling}.

To date, the atipicallity of satellite distributions can only be
addressed in simulations.
Most of the studies have been performed in the currently preferred
paradigm, which is based on a Cold Dark Matter cosmology in a expanding
Universe described by General Relativity including a Cosmological
Constant, the so-called Lambda Cold Dark Matter (LCDM) model.
The broad consensus is that these highly structured satellite distributions
are hard to find in simulations.  
There are still disagreements on the precise degree of atypicallity
and the conditions required to simulate satellite distributions closer.
These disagreements seem to be originated on what is considered to be
a LG analog in the simulation.

Some studies have focused on the study of individual dark matter halos
that could host galaxies like the MW or M31, while some other studies
have tried to simulate the formation of galaxy pairs resembling both
the MW and M31.
The transition from individual halos to pairs has been
motivated by studies that show how the LG location in the cosmic web
plays a role in distingushing the the MW and M31 halos from the
average $\sim 10^{12}$ \Msun\ DM halo \citep{2011MNRAS.417.1434F}
Studies of the LG as a pair of isolated DM halos have also found that
the LG itself has relatively atypical kinematicsin the context
of LCDM and a strong preference to lie along filaments \citep{ForeroRomero2013}.
Putting together these pieces, the cosmic web should therefore play a
determinant role in building up planar satellite distributions through
preferential alignments and accretions histories of galaxies
\citep{2014MNRAS.443.1090F}.

The aim of our work is to overcome two difficulties in the study of
atipicallity in satellite distributions.
First, atypicallity is usually computed from distributions comming from
different simulated pair samples, making it inconvenient to compare
results from different studies. 
Second, halo pairs with similar to the LG are scarse in cosmological
volumes, making it hard to make direct comparisons against simulated
systems that fully resemble the LG.

We surmount the first difficulty by setting as a point of reference an
spherical satellite distribution and not LCDM simulations. 
This spherical distribution is built from the data itself
(observational or simulated) by randomizing its angular position
around the central galaxy and keeping fixed its radial distance
\citep{2017AN....338..854P}. 
We characterize each satellite distribution with its inertia
tensor and summarize the results with the scalars describing the width
of the distribution and their axis ratios.

@@

In Section we list the sources of the observational and
simulated data to be used throughout the paper.
In Section we describe the methods we use to build halo pairs and 
characterize its satellite distributions.
In Section we present our results to finally present our conclusions
in Section.

\section{Data samples}

@@

The information in the catalog is based on the catalogue compiled by
\cite{2012AJ....144....4M}.
The distance estimate are based on resolved stellar populations. 
We use three dimensional positions in a cartesian Coordinate system
as computed by \cite{2013MNRAS.435.1928P}.
In this coordinate system the $z$-axis points toward the Galactic north pole, the
$x-$-axis points in the direction from the Sun to the Galactic centre,
and the $y$-axis points in the direction of the Galactic rotation.

@@

We buid a sample of pairs that resemble the conditions in the LG.
To construct this sample we select from Illustris-1 all galaxies with an stellar
mass in the range $1\times10^{10}\Msun <M_{\star}<1.5 \times 10^{11}
\Msun$.
Then we select the pairs with the following conditions.

@@

We find 27 pairs with these conditions. 
We then select the pairs where in both halos there are at least 15
detected subhalos, thus discarding pairs with halos with the lowest
mass.
We end up with a total of 20 pairs in Illustris-1.
In Illustris-1-DM we use the center of mass position of the 27 pairs
in Illustris-1 to find the matching halo pairs.
After discarding the pairs with less than 15 detected subhalos in one
of the halos we end up with a toal of 24 pairs. 
This corresponds to a pair number density of $\sim 2 \times10^{-5}$
pairs Mpc$^{-3}$ 
Appendix A shows the physical  properties (stellar masses, maximum
circular velocities, radial velocities and separation) in those
pairs.

Although Illustris-1 has stellar particles, we do not use their
properties to select the satelite population because the smallest
galaxies are barely resolved in stellar mass at magnitudes of
$M_V=-9$, close to the limit of the 11 ``classical'' MW satellites.
We use instead the dark matter information as the smallest sub-halos
we care about are sampled with at least $40$ particles.

Both in Illustris-1 and Illustris-1-Dm we chose the satellite samples
by ranking the subhalos in decreasing order of its current maximum
circular velocity and select the first $N_p$ halos in the list.  
The results presented here correspond to $11\leq N_p\leq 15$.

@@

We use data from the public release of the Exploring the Local
Universe In Simulations (ELVIS) project.
For a detailed description of that project and its data we refer the
reader to \cite{XX}. 
Here we summarize the elements relevant to our discussion.

@@

This corresponds to a pair number density of $\sim 8 \times10^{-6}$ pairs
Mpc$^{-3}$, this is a factor $\sim 2.5$ lower than the pair number density
we find in the Illustris data.
There were a total of 146 pairs that met those criteria, but only $12$
were chosen for resimulation. 
Additionally, the selected pairs for resimulation have a relative
tangential velocity less than $75$ \kms. 
The dark matter particle resolution in these resimulations is
$1.9\times 10^5$, a thousand times larger than Illustris-1.
In this paper we only use the results from these $12$ resimulated pairs.

@@

We compute the satellite statistics fo 11 up to 15 satellites.
The lowest bound corresponds to the number of classical Milky Way
satellites.
The upper limit corresponds to the maximum number of satellites that
can be resolved in both halos for most of the isolated pairs in Illustris-1.
In simulations we rank the subhalos by their maximum
circular velocity, in observations we rank the satellites by its $M_V$
magnitude.

@@

\begin{table*}
  \centering
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{|p{2.5cm}|c|c|c|c|c|c|}
    \hline
    \multirow{2}{4.0cm}{} & \multicolumn{2}{c|}{\textbf{Observations}}
    & \multicolumn{2}{c|}{\textbf{Randomized Obs.}} &
    \multicolumn{2}{c|}{\textbf{Normalized Units}}\\
    % \hline
    % \textbf{Inactive Modes} & \textbf{Description}\\
    \cline{2-7}
    & \textbf{M31} & \textbf{MW} & \textbf{M31} & \textbf{MW} & \textbf{M31} & \textbf{MW} \\
    %\hhline{~--}
    \hline
    Plane width (kpc) & $59\pm 3$  & $21\pm 2$  & $65\pm 12$   &
    $45\pm 8$    & $-0.48\pm 0.24$ & $-2.48\pm 0.26$\\\hline
    $c/a$ ratio & $0.45\pm 0.04$ & $0.25\pm 0.05$ & $0.55\pm0.10$ &
    $0.53\pm 0.10$ & $-1.03\pm 0.37$ & $-2.18\pm 0.42$\\ \hline
    $b/a$ ratio & $0.82\pm 0.06$ & $0.80\pm 0.04$ & $0.82\pm0.07$ &
    $0.81\pm 0.08$ & $-0.02\pm 0.82$ & $-0.13\pm 0.47$\\ \hline
  \end{tabular}
  \caption{Results from observations. Mean values and standard deviations for the different
    quantities describing satellite distribution: plane width, $c/a$
    ratio and $b/a$ ratio. 
    The first colum refers to the results from
    observational data, the second column uses the spherically
    randomized version of the observational data, the third column
    corresponds to the observational data recentered and normalized
    by the randomized values. \label{table:observations}}
\end{table*}

\begin{table*}
  \centering
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{|p{2.5cm}|c|c|c|c|c|c|}
    \hline
    \multirow{2}{4.0cm}{} & \multicolumn{2}{c|}{\textbf{Illustris-1-Dark}} & \multicolumn{2}{c|}{\textbf{Illustris-1}} & \multicolumn{2}{c|}{\textbf{ELVIS}}\\
    % \hline
    % \textbf{Inactive Modes} & \textbf{Description}\\
    \cline{2-7}
    & \textbf{M31} & \textbf{MW} & \textbf{M31} & \textbf{MW}& \textbf{M31} & \textbf{MW}\\
    %\hhline{~--}
    \hline
    Plane width (kpc) & $62\pm 3$   & $61\pm 3$     & $70\pm 4$ & $67\pm 2$ & $70\pm 2$& $68\pm 4$ \\\hline
    $c/a$ ratio & $0.50\pm0.01$ & $0.50\pm 0.02$ & $0.52\pm 0.01$ & $0.53\pm 0.01$ & $0.54\pm 0.01$& $0.49\pm 0.02$ \\ \hline
    $b/a$ ratio & $0.79\pm0.01$ & $0.81\pm 0.02$ & $0.80\pm 0.01$ & $0.80\pm 0.02$ & $0.80\pm0.01$& $0.81\pm 0.01$\\ \hline
  \end{tabular}
  \caption{Results from simulations. Mean values and standard deviations for the different
    quantities describing satellite distributions: plane width, $c/a$
    ratio and $b/a$ ratio. 
    The first column summarize the results from the 24 pairs in
    Illustris-1-Dark, the second column from the 20 pairs in
    Illustris-1 and the third column from the 12 pairs in the
    ELVIS project.\label{table:simulations}}
\end{table*}

\subsection{Describing Samples with the Inertia Tensor}
We base all our results on the description provided by the inertia
tensor defined by the satellites's positions.

\begin{equation}
{\bf{\bar{I}}} = \sum_{k=1}^{N_s}[(\bf{r}_i - \bf{r}_0)^2\cdot \bf{1} -
  (\bf{r}_i-\bf{r}_0)\cdot (\bf{r}_i - \bf{r}_0)^{T}],
\end{equation}
%
where $k$ indexes the set of satellites of interest
$\bf{r}_k$ are the satellites' positions, $\bf{r}_{0}$ is the location
of the central galaxy $\bf{1}$ is the unit matrix,  and  
${\bf r}^T$ is the transposed vector $\bf{r}$. 
We use $\bf{r}_0$ as the position of the central galaxy, and not the
satellites' geometrical center, to allow for a fair comparison once
the angular positions of the satellites are randomized around this
point.

@@

We compare every satellite distribution against its own spherically
randomized distribution.
We keep fixed the radial position of every satellite
with respect to the central galaxy and then randomizing its angular
position. 
We repeat this procedure 1000 times for each satellite distribution
and proceed to measure the quantities mentioned in the last section:
$w$, $c/a$ and $b/a$.
For each quantity we compute an average and standard deviation from
the 1000 random samples. 
This allows us to build a normalized version of all quantities of
interest by substracting the mean and dividing between the standard
deviation of the randomized samples.

This allows us to make a first comparison. 
The observed/simulated
distribution against its randomized version. 
The second intermediate comparison we make are observations against
simulations.
The final comparison is between the normalized quantities, both
observed and simulated.

The comparison between the normalized quantities is the one that
carries the important information about the deviations from
spheriticy. 
We do not want to directly compare how the results observations
deviate from simulations but to compare the deviations from
asphericity in observations and simulations.

@@

We compute the preferred covariance matrix and the mean distribution values
with a jackknife technique. 
That is, out of the the $n$ pairs we have, we perfom $n$ different
covariance and mean value measurement using only $n-1$ pairs. 
The reported covariance and mean values correspond to the average of
all measurements, the corresponding standard deviation also help us to
estimate the uncertainty on every reported coefficient.
This compact description allows us to generate samples of size $N$
that are consistent by construction with their parent simulation.

@@

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_width.pdf}
\caption{Plane width characterization in observations and
  simulations (Illustris1-Dark in this case.) In all panels the
  horizontal/vertical axis corresponds to the  M31/MW or the
  most/less massive halo in the pair 
  Left. Plane width in physical units comparing the results from observations
(stars) against the result of spherically randomizing the satellite
positions (circles). 
Middle. Average from the observations (star) and the average from each
pair in the simulation (circles).
Right. Same as the middle panel only that this time each point has
been normalized (median substracted and normalized by the standar
deviation) to the results of its randomization. 
The main message of these panels that the MW has a significantly
thiner plane both compared to the result of its own satellite
spherical randomization (left panel) and the expectation from
simulations (middle and right panel).  
This low value is $2\sigma$ away from what is expected in a spherical
distribution. 
In M31 the satellite distribution are in agreement with the
expectations both from an spherical distribution and the results form the
simulations. 
\label{fig:scatter_width}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_ca_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_ca_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_ca_ratio.pdf}
\caption{Same layout as in Figure \ref{fig:scatter_width}. 
This time for the $c/a$ axis ratio. 
The same message holds in this case; namely, that 
the MW is has a significantly lowwe $c/a$ value than the
expectation from the spherical distribution and the simulations. 
This low value is also $2\sigma$ away from the expactations for an
spherical distribution.
On the other hand, M31 is consistent both with an spherical
distribution and the results from simulations.
\label{fig:scatter_ca_ratio}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_ba_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_ba_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_ba_ratio.pdf}
\caption{Same layout as in Figure \ref{fig:scatter_width}. 
This time for the $b/a$ axis ratio. In this case both the MW and M31
are consistent with the results of a spherical distribution and 
simulations. 
\label{fig:scatter_ba_ratio}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.45\textwidth]{gaussian_model_illustrisdm_M31.pdf}
\includegraphics[width=0.45\textwidth]{gaussian_model_illustrisdm_MW.pdf}
\caption{Results from multivariate gaussian model fitted on   the
  normalized values for the plane width $w$, $c/a$ ratio and $b/a$
  ratio from the Illustris1-Dark results.  
  Left/right panel correspond
  to M31/MW.  
  The cross indicates the observed values.
  The contour levels in the 2D histograms correspond to the $1\sigma$,
$2\sigma$ and $3\sigma$ contours in two dimensions.
  The dahsed vertical lines in the histograms along the diagonal
  correspond to the $1\sigma$ boundaries in one dimension.
  The results for the gaussian model are built from $10^6$ point
  realizations in the three-dimensional space spaned by the variables of
  interest for each halo. 
  This plot clearly shows how the M31 results are well within the
  expectations from simulations while the MW has an unusual low value for
  the plane width and the $c/a$ axis ratio.
\label{fig:correlations_illustrisdm}}
\end{figure*}

@@

Table \ref{table:observations} and Table \ref{table:simulations}
summarize the mean values and uncertainties for the plane $w$, $c/a$
ratio and $b/a$ ratio in the observations and simulations, respectively. 
The uncertainty in the observations is computed over the results with
different number of satellites, while in the simulations it
corresponds to the standard deviation over different halos.

@@

\section{Conclusions}

>>>>

\begin{abstract}
We use numerical simulations to build an explicit probability
distribution for the asphericity in the satellite distribution around
galaxies similar to the Local Group (LG) in the Lambda Cold Dark
Matter (LCDM) paradigm. 
This allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that fully resemble the LG.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-Dark and ELVIS. 
Detailed results differ among the simulations suggesting a strong
influence of the typical Dark Matter (DM) halo mass in the different pair samples.  
However, there are three common trends in these simulations. 
First, at most of $2\%$ the pairs are expected to have satellite
distributions with the same asphericity as the LG; second,
between $27\%$ to $56\%$ of the pairs have a halo with a satellite
distribution as aspherical as in M31; and third, at most $3\%$ of the
pairs have a satellite distribution as planar as in the MW. 
The brightest satellites in the M31 galaxy turn out to have a rather
typical distribution in the LCDM context, while the MW distribution is atypical in
this respect.  
These results place the LG at the level of a $3\sigma$ outlier
in the LCDM paradigm. 
The approach presented here facilitates the comparison between
different numerical setups and choices to select halo pairs in simulations.
Quantifying the probability density for the satellite distribuition
asphericity as a function of halo mass and large scale environment 
will be a necessary step to understand the origin of the LG
atipicallity in the LCDM context as presented here.  
\end{abstract}

@@

The spatial distribution of satellite galaxies around our Milky Way
(MW) and the M31 galaxy provides a stringent test for structure
formation theories in a explicit cosmological context
\citep{2015ApJ...815...19P, 2016MNRAS.457.1931S, 2016arXiv160601516L}.  
The presence of a Vast Polar Structure of satellites around the Milky
Way has been stablished in the last decade \citep{2013MNRAS.435.1928P}.
This structure is present in the $11$ brightest "classical" dwarf
galaxies and has been confirmed as fainter satellites have been
detected \citep{2014MNRAS.441.2124B}. %XXXXXXX  
In the case of M31, observational studies have found a planar satellite
distribution in a subset of $15$ satellites out of the 
%homogeneous parent sample
%comentario: yo no describiría la poblacion como homogenea porque la
%mayoría están entre M31 y la MW la poblción está "loopsided"
total poulation of $27$ satellites.\citep{2013ApJ...766..120C,
2013Natur.493...62I} 
However, taking into account only the $11$ brightest satellites (as a
comparison with the MW) there is no evidence of any planar structure
in M31.

The question whether this kind of planar configurations are atypical
in the observable Universe cannot be properly addressed given the
current observational limitations. %citation?
Comparisons against observations in large galaxy surveys have only
been able to constrain that $5\%$ of the Milky-Way like galaxies in
the Universe have two bright satellite galaxies as is the case of the
Milky Way and its Large and Small Magellanic Clouds.%citation 
A planar structure ouside the Local Group (LG) has only been confirmed
in Centaurus A \citep{muller2018whirling}.

To date, the atipicallity of satellite distributions can only be
addressed in simulations.
Most of the studies have been performed in the currently preferred
paradigm, which is based on a Cold Dark Matter cosmology in a expanding
Universe described by General Relativity including a Cosmological
Constant, the so-called Lambda Cold Dark Matter (LCDM) model.
The broad consensus is that these highly structured satellite distributions
are hard to find in simulations \citep{}.  
There are still disagreements on the precise degree of atypicallity
and the conditions required to simulate satellite distributions closer.
These disagreements seem to be originated on what is considered to be
a LG analog in the simulation.

Some studies have focused on the individual dark matter halos
that could host galaxies like the MW or M31, while some other studies
have tried to simulate the formation of galaxy pairs resembling both
the MW and M31.
The transition from individual halos to pairs has been
motivated by studies that show how the LG location in the cosmic web
plays a role in distinguishing the MW and M31 halos from the
average $\sim 10^{12}$ \Msun\ DM halo \citep{2011MNRAS.417.1434F}
Studies of the LG as a pair of isolated DM halos have also found that
the LG itself has relatively atypical kinematics in the context
of LCDM and a strong preference to lie along filaments \citep{ForeroRomero2013}.
Putting together these pieces, the cosmic web should therefore play a
determinant role in building up planar satellite distributions through
preferential alignments and accretions histories of galaxies
\citep{2014MNRAS.443.1090F}.

The aim of our work is to overcome two difficulties in the study of
atipicallity in satellite distributions.
First, atypicallity is usually computed from distributions comming from
different simulated pair samples, making it inconvenient to compare
results from different studies. 
Second, halo pairs similar to the LG are scarse in cosmological
volumes, making it hard to perform direct comparisons against simulated
systems that fully resemble the LG.

We surmount the first difficulty by setting as a point of reference a
spherical satellite distribution and not LCDM simulations.
%comentario: esta explicación no me parece muy clara
This spherical distribution is built from the data itself
(observational or simulated) by randomizing its angular position
around the central galaxy and keeping its radial distance fixed
\citep{2017AN....338..854P}. 
We characterize each satellite distribution with its inertia
tensor and summarize the results with the scalars describing the width
of the distribution and its axis ratios.

@@

In Section \ref{sec:DataSamples} we list the sources of the observational and
simulated data to be used throughout the paper.
In Section \ref{sec:SpatialMeasurements} we describe the methods we use to build halo pairs and 
characterize its satellite distributions.
In Section \ref{sec:results} we present our results to finally
conclude in Section \ref{sec:conclusions}.

\section{Data samples}\label{sec:DataSamples}

@@

The information in the catalogue is based on the catalogue compiled by
\cite{2012AJ....144....4M}.
The distance estimates are based on resolved stellar populations. 
We use three dimensional positions in a cartesian coordinate system
as computed by \cite{2013MNRAS.435.1928P}.
In this coordinate system the $z$-axis points towards the Galactic north pole, the
$x$-axis points in the direction from the Sun to the Galactic centre,
and the $y$-axis points in the direction of the Galactic rotation.

@@

We buid a sample of pairs that resemble the conditions in the LG.
To construct this sample we select from Illustris-1 all galaxies with a stellar
mass in the range $1\times10^{10}\Msun <M_{\star}<1.5 \times 10^{11}
\Msun$.
Then we select the pairs with the following conditions.

@@

We find 27 pairs with these conditions. 
We then select the pairs where in both halos there are at least 15
detected subhalos, thus discarding pairs with halos with the lowest
mass.
We end up with a total of 20 pairs in Illustris-1.
In Illustris-1-DM we use the center of mass position of the 27 pairs
in Illustris-1 to find the matching halo pairs.
After discarding the pairs with less than 15 detected subhalos in one
of the halos we end up with a total of 24 pairs. 
This corresponds to a pair number density of $\sim 2 \times10^{-5}$
pairs Mpc$^{-3}$ 
Appendix A shows the physical  properties (stellar masses, maximum
circular velocities, radial velocities and separation) of those
pairs.

Although Illustris-1 has stellar particles, we do not use their
properties to select the satelite population because the smallest
galaxies are barely resolved in stellar mass at magnitudes of
$M_V=-9$, close to the limit of the 11 ``classical'' MW satellites.
We use instead the dark matter information as the smallest sub-halos
we care about are sampled with at least $40$ particles.  
%comentario: este último criterio no me parece que esté claramente
%descrito

Both in Illustris-1 and Illustris-1-DM we chose the satellite samples
by ranking the subhalos in decreasing order of its current maximum
circular velocity and select the first $N_p$ halos in the list.  
The results presented here correspond to $11\leq N_p\leq 15$.

@@

We use data from the public release of the Exploring the Local
Universe In Simulations (ELVIS) project.
For a detailed description of that project and its data we refer the
reader to \cite{2014MNRAS.438.2578G}. 
Here we summarize the elements relevant to our discussion.

@@

This corresponds to a pair number density of $\sim 8 \times10^{-6}$ pairs
Mpc$^{-3}$, this is a factor $\sim 2.5$ lower than the pair number density
we find in the Illustris data.
There were a total of 146 pairs that met those criteria, but only $12$
were chosen for resimulation. 
Additionally, the selected pairs for resimulation have a relative
tangential velocity less than $75 $\kms. 
The dark matter particle resolution in these resimulations is
$1.9\times 10^5$, a thousand times larger than Illustris-1.
In this paper we only use the results from these $12$ resimulated pairs.

@@

We compute the satellite statistics for 11 up to 15 satellites.
The lower limit corresponds to the number of classical Milky Way
satellites.
The upper limit corresponds to the maximum number of satellites that
can be resolved in both halos for most of the isolated pairs in Illustris-1.
In simulations we rank the subhalos by their maximum
circular velocity, in observations we rank the satellites by its $M_V$
magnitude.

@@

\begin{table*}
  \centering
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{|p{2.5cm}|c|c|c|c|c|c|}
    \hline
    \multirow{2}{4.0cm}{} & \multicolumn{2}{c|}{\textbf{Observations}}
    & \multicolumn{2}{c|}{\textbf{Randomized Obs.}} &
    \multicolumn{2}{c|}{\textbf{Normalized Units}}\\
    % \hline
    % \textbf{Inactive Modes} & \textbf{Description}\\
    \cline{2-7}
    & \textbf{M31} & \textbf{MW} & \textbf{M31} & \textbf{MW} & \textbf{M31} & \textbf{MW} \\
    %\hhline{~--}
    \hline
    Plane width (kpc) & $59\pm 3$  & $21\pm 2$  & $65\pm 12$   &
    $45\pm 8$    & $-0.48\pm 0.24$ & $-2.48\pm 0.26$\\\hline
    $c/a$ ratio & $0.45\pm 0.04$ & $0.25\pm 0.05$ & $0.55\pm0.10$ &
    $0.53\pm 0.10$ & $-1.03\pm 0.37$ & $-2.18\pm 0.42$\\ \hline
    $b/a$ ratio & $0.82\pm 0.06$ & $0.80\pm 0.04$ & $0.82\pm0.07$ &
    $0.81\pm 0.08$ & $-0.02\pm 0.82$ & $-0.13\pm 0.47$\\ \hline
  \end{tabular}
  \caption{Results from observations. Mean values and standard deviations for the different
    quantities describing the satellite distributions: plane width, $c/a$
    ratio and $b/a$ ratio. 
    The first colum refers to the results from
    observational data, the second column uses the spherically
    randomized version of the observational data, the third column
    corresponds to the observational data recentered and normalized
    by the randomized values. \label{table:observations}}
\end{table*}

\begin{table*}
  \centering
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{|p{2.5cm}|c|c|c|c|c|c|}
    \hline
    \multirow{2}{4.0cm}{} & \multicolumn{2}{c|}{\textbf{Illustris-1-Dark}} & \multicolumn{2}{c|}{\textbf{Illustris-1}} & \multicolumn{2}{c|}{\textbf{ELVIS}}\\
    % \hline
    % \textbf{Inactive Modes} & \textbf{Description}\\
    \cline{2-7}
    & \textbf{M31} & \textbf{MW} & \textbf{M31} & \textbf{MW}& \textbf{M31} & \textbf{MW}\\
    %\hhline{~--}
    \hline
    Plane width (kpc) & $62\pm 3$   & $61\pm 3$     & $70\pm 4$ & $67\pm 2$ & $70\pm 2$& $68\pm 4$ \\\hline
    $c/a$ ratio & $0.50\pm0.01$ & $0.50\pm 0.02$ & $0.52\pm 0.01$ & $0.53\pm 0.01$ & $0.54\pm 0.01$& $0.49\pm 0.02$ \\ \hline
    $b/a$ ratio & $0.79\pm0.01$ & $0.81\pm 0.02$ & $0.80\pm 0.01$ & $0.80\pm 0.02$ & $0.80\pm0.01$& $0.81\pm 0.01$\\ \hline
  \end{tabular}
  \caption{Results from simulations. Mean values and standard deviations for the different
    quantities describing the satellite distributions: plane width, $c/a$
    ratio and $b/a$ ratio. 
    The first column summarize the results from the 24 pairs in
    Illustris-1-Dark, the second column from the 20 pairs in
    Illustris-1 and the third column from the 12 pairs in the
    ELVIS project.\label{table:simulations}}
\end{table*}

\subsection{Describing Samples with the Inertia Tensor}
We base all our results on the description provided by the inertia
tensor defined by the satellites' positions.

\begin{equation}
{\bf{\bar{I}}} = \sum_{k=1}^{N_s}[(\bf{r}_i - \bf{r}_0)^2\cdot \bf{1} -
  (\bf{r}_i-\bf{r}_0)\cdot (\bf{r}_i - \bf{r}_0)^{T}],
\label{eq:intensor}
\end{equation}
%
where $k$ indexes the set of satellites of interest
$\bf{r}_k$ are the satellites' positions, $\bf{r}_{0}$ is the location
of the central galaxy $\bf{1}$ is the unit matrix, and  
${\bf r}^T$ is the transposed vector $\bf{r}$. 
We use $\bf{r}_0$ as the position of the central galaxy, and not the
satellites' geometrical center, to allow for a fair comparison once
the angular positions of the satellites are randomized around this
point.

@@

We compare every satellite distribution against its own spherically
randomized distribution.
We keep fixed the radial position of every satellite
with respect to the central galaxy and then randomize its angular
position. 
We repeat this procedure 1000 times for each satellite distribution
and proceed to measure the quantities mentioned in the previous section:
$w$, $c/a$ and $b/a$.
For each quantity we compute the average and standard deviation from
the 1000 random samples. 
This allows us to build a normalized version of all quantities of
interest by substracting the mean and dividing between the standard
deviation of the randomized samples.

This allows us to make a first comparison. 
The observed/simulated
distribution against its randomized version. 
The second intermediate comparison we make is of the observations against
the simulations.
The final comparison is between the normalized quantities, both
observed and simulated.

The comparison between the normalized quantities is the one that
carries the important information about the deviations from
spheriticy. 
We do not want to directly compare how the observations
deviate from simulations but to compare the deviations from
asphericity in observations and simulations.

@@

We compute the preferred covariance matrix and the mean distribution values
with a jackknife technique. 
That is, out of the the $n$ pairs we have, we perfom $n$ different
covariance and mean value measurements using only $n-1$ pairs. 
The reported covariance and mean values correspond to the average of
all measurements, the corresponding standard deviation also helps us to
estimate the uncertainty on every reported coefficient.
This compact description allows us to generate samples of size $N$
that are consistent by construction with their parent simulation.

@@

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_width.pdf}
\caption{Plane width characterization in observations and
  simulations (Illustris1-Dark in this case). In all panels the
  horizontal/vertical axis corresponds to the  M31/MW or the
  most/less massive halo in the pair 
  Left. Plane width in physical units comparing the results from observations
(stars) against the result of spherically randomizing the satellite
positions (circles). 
Middle. Average from the observations (star) and the average from each
pair in the simulation (circles).
Right. Same as the middle panel except that this time each point has
been normalized (median substracted and normalized by the standard
deviation) to the results of its randomization. 
The main message of these panels is that the MW has a significantly
thinner plane both compared to the result of its own satellite
spherical randomization (left panel) and the expectation from
simulations (middle and right panel).  
This low value is $2\sigma$ away from what is expected in a spherical
distribution. 
In M31 the satellite distribution is in agreement with the
expectations both from an spherical distribution and the results form the
simulations. 
\label{fig:scatter_width}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_ca_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_ca_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_ca_ratio.pdf}
\caption{Same layout as in Figure \ref{fig:scatter_width}. 
This time for the $c/a$ axis ratio. 
The same message holds in this case; namely, that 
the MW has a significantly lower $c/a$ value than the
expectation from the spherical distribution and from the simulations. 
This low value is also $2\sigma$ away from the expactations for a
spherical distribution.
On the other hand, M31 is consistent both with a spherical
distribution and with the results from simulations.
\label{fig:scatter_ca_ratio}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_ba_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_ba_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_ba_ratio.pdf}
\caption{Same layout as in Figure \ref{fig:scatter_width}. 
This time for the $b/a$ axis ratio. In this case both the MW and M31
are consistent with the results of a spherical distribution and 
the simulations. 
\label{fig:scatter_ba_ratio}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.45\textwidth]{gaussian_model_illustrisdm_M31.pdf}
\includegraphics[width=0.45\textwidth]{gaussian_model_illustrisdm_MW.pdf}
\caption{Results from multivariate gaussian model fitted to the
  normalized values for the plane width $w$, $c/a$ ratio and $b/a$
  ratio from the Illustris1-Dark results.  
  Left/right panel correspond to M31/MW respectively.  
  The cross indicates the observed values.
  The contour levels in the 2D histograms correspond to the $1\sigma$,
$2\sigma$ and $3\sigma$ contours in two dimensions.
  The dashed vertical lines in the histograms along the diagonal
  correspond to the $1\sigma$ boundaries in one dimension.
  The results for the gaussian model are built from $10^6$ point
  realizations in the three-dimensional space spaned by the variables of
  interest for each halo. 
  This plot clearly shows how the M31 results are well within the
  expectations from simulations while the MW has an unusual low value for
  the plane width and the $c/a$ axis ratio.
\label{fig:correlations_illustrisdm}}
\end{figure*}

@@

Table \ref{table:observations} and Table \ref{table:simulations}
summarize the mean values and uncertainties for the plane $w$, $c/a$
ratio and $b/a$ ratio in the observations and simulations, respectively. 
The uncertainty in the observations is computed from the results with
different number of satellites, while in the simulations it
corresponds to the standard deviation over different halos.

@@

\section{Conclusions}\label{sec:conclusions}
